---
title: "Robustness & Bias in NLP"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

This notebook contains code for producing the table in the paper. 

```{r}
library(tidyverse)
library(kableExtra)
library(janitor)

source("utils-R/utility.R")
```

## Read in File, Average Values ! 
```{r read files}
perf <- readbulk::read_bulk(directory = "results", extension = ".csv")

perf = perf %>% 
  mutate(across(ends_with("_acc"), ~ .x * 100),
         across(ends_with("_f"), ~ .x * 100),
         across(ends_with("_p"), ~ .x * 100),
         across(ends_with("_r"), ~ .x * 100),
         across(starts_with("dep"), ~ .x * 100)
         )
```

## Define Scoring Function (Pairwise Bonferroni T-tests)
```{r scoring summary functions}
# t-test function for the score_to_df function
score_t_test = function(mdl, augmenter, default, data, score){
  d = data %>% filter(model == mdl)

  x = d[[score]][d$augmenter == augmenter]
  mu = d[[score]][d$augmenter == default]
  
  if (length(mu) == 1){
    t = t.test(x = x,
           mu = mu, paired = FALSE, var.equal = FALSE,
           conf.level = 0.95)
  } else {
    t = t.test(x = x,
           y = mu, paired = FALSE, var.equal = FALSE,
           conf.level = 0.95) 
  }
  return (t)
}

# score_to_df function: bonferroni adjusted t-tests for all augmentations for all models against a baseline augmentation
score_to_df = function(data, score, baseline){
  dfs = NULL
  i = 1
  for(mdl in unique(data$model)){
    for (aug in unique(data$augmenter)){
      v = data[[score]][data$augmenter == aug & data$model == mdl]
      
      if (length(v) <= 2){
          dfs[[i]] = tibble(model=mdl, augmenter=aug, mean=v[1], sd=NA, 
                    conf_int = "",
                    p_value=NA)
          i = i+1
          next
      }
      
      mu = mean(v)
      sigma = sd(v)
      print(paste(mdl, aug, length(v)))
      t = score_t_test(mdl=mdl, augmenter=aug, default=baseline, data=data, score=score)
      
      p = p.adjust(t$p.value, method = "bonferroni", n = 6)

      dfs[[i]] = tibble(model=mdl, augmenter=aug, mean=mu, sd=sigma, 
                        conf_int = paste("(",  round(t$conf.int[1], 2),", ",  round(t$conf.int[2], 2), ")", sep = ""),
                        p_value=p)
      i = i+1
    }
  }
  return(bind_rows(dfs))
}
```

## Define function for calculating and creating table for chosen value 
```{r}
create_metric_table = function(metric_col_name, data, metric_label, order_groups, order_models){
  # create metric df 
  metric_df <- data %>% 
  select(model, 
         metric_col_name,
         augmenter = augmenter
  ) %>% 
  filter(augmenter != "Input size augmentation 10 sentences" & augmenter != "Input size augmentation 5 sentences")

  # calculate scores 
  metric_df_scores <- score_to_df(metric_df, metric_col_name, "Danish names")
  
  # create performance table with p-values, standard deviation in parenthesis
  metric_table = metric_df_scores %>% arrange(factor(model, levels = order_models)) %>% 
    mutate(mean = paste(format(round_half_up(mean, 1), digits = 1, nsmall=1)),
           p_value_star = if_else(p_value < 0.05, "*", "", missing =""),
           string_value = if_else(is.na(sd), paste(mean, sep=""),
                                  paste(mean,  " (", 
                                        format(round_half_up(sd, 1), digits = 1, nsmall=1), 
                                        ")", p_value_star, sep="")),
           model = plyr::mapvalues(model, from=names_from, to=names_to),
           augmenter = plyr::mapvalues(augmenter, from=aug_names_from, to=aug_names_to)) %>% 
    select(-c(mean, sd, p_value, p_value_star, conf_int)) %>% 
    pivot_wider(names_from = augmenter, values_from=c(string_value)) %>% 
    select(Model=model, all_of(aug_names_to)) # reorder
  
  # add metric col, specifying the metric 
  Metric <- rep(metric_label,11)
  metric_table <- cbind(metric_table, Metric)
  
  # re order columns 
  metric_table <- metric_table[, c("Model", "Metric", order_groups)]

  return(metric_table)
}
```

### Define function for collapsing model names 
```{r}
# collapse rows
collapse_rows_df <- function(df, variable){
  group_var <- enquo(variable)
  df %>%
    group_by(!! group_var) %>%
    mutate(groupRow = 1:n()) %>%
    ungroup() %>%
    mutate(!!quo_name(group_var) := ifelse(groupRow == 1, as.character(!! group_var), "")) %>%
    select(-c(groupRow))
}
```
Function credits: https://stackoverflow.com/questions/51450402/how-to-collapse-groups-of-row-values-within-a-table-using-formattable 

## Rename model and augmenters
```{r rename model and aug}
# original model names
names_from = c("spacy_large", "spacy_medium", "dacy_large", "dacy_medium", "danlp_bert", "spacy_small", "flair", "nerda_bert", "dacy_small", "polyglot", "scandi_ner")

# new model names
names_to = c("SpaCy large", "SpaCy medium", "DaCy large", "DaCy medium", "DaNLP BERT", "SpaCy small", "Flair", "NERDA", "DaCy small", "Polyglot", "ScandiNER")

# aug names original
aug_names_from = c("Danish names", "Muslim names", "Female names", "Male names", "Muslim female names","Muslim male names", "Unisex names") 

# aug names new 
aug_names_to = c("Majority all", "Minority all", "Majority Women", "Majority Men", "Minority Women", "Minority Men", "Unisex")
```

## Tables 1 and 2!  
## Define order of models and groups
```{r}
order_models <- c("scandi_ner","dacy_large", "dacy_medium", "dacy_small", "danlp_bert", "flair", "nerda_bert", "spacy_large", "spacy_medium", "spacy_small", "polyglot")

order_groups = c("Majority all", "Minority all", "Majority Men", "Minority Men", "Majority Women", "Minority Women", "Unisex")
```


### Table 1: Precision and Recall: NER without MISC 
#### Calculate precision and recall
```{r}
ent_per_tbl_precision_tb1 <- create_metric_table("ents_excl_MISC_ents_p", perf, "Precision", order_groups, order_models)
ent_per_tbl_recall_tb1 <- create_metric_table("ents_excl_MISC_ents_r", perf, "Recall", order_groups, order_models)
```

#### Combine two dataframes 
```{r}
both_metrics_tb1 <- rbind(ent_per_tbl_precision_tb1, ent_per_tbl_recall_tb1)

# define new order with the actual names 
new_order <- c("ScandiNER", "DaCy large", "DaCy medium", "DaCy small", "DaNLP BERT", "Flair", "NERDA", "SpaCy large", "SpaCy medium", "SpaCy small", "Polyglot")

# reorder so that precision and recall comes after each other 
both_metrics_tb1 <- both_metrics_tb1[order(unlist(sapply(both_metrics_tb1$Model, function(x) which(new_order == x)))),] 

# reset index 
rownames(both_metrics_tb1) <- NULL 
```

### Table 2 Precision and Recall: PER score  
#### Calculate precision and recall
```{r}
ent_per_tbl_precision_tb2 <- create_metric_table("ents_per_type_PER_p", perf, "Precision", order_groups, order_models)
ent_per_tbl_recall_tb2 <- create_metric_table("ents_per_type_PER_r", perf, "Recall", order_groups, order_models)
```

#### Combine two dataframes 
```{r}
both_metrics_tb2 <- rbind(ent_per_tbl_precision_tb2, ent_per_tbl_recall_tb2)

# define new order with the actual names 
new_order <- c("ScandiNER", "DaCy large", "DaCy medium", "DaCy small", "DaNLP BERT", "Flair", "NERDA", "SpaCy large", "SpaCy medium", "SpaCy small", "Polyglot")

# reorder so that precision and recall comes after each other 
both_metrics_tb2 <- both_metrics_tb2[order(unlist(sapply(both_metrics_tb2$Model, function(x) which(new_order == x)))),] 

# reset index 
rownames(both_metrics_tb2) <- NULL 
```

## Make Tables 
### Table 1: 
```{r}
collapse_rows_df(both_metrics_tb1, Model) %>% 
  kbl(
    booktabs=T, 
    col.names = c("Model", "Metric", "Majority", "Minority", "Majority", "Minority", "Majority", "Minority", "Majority"),
    caption = "NER Performance (Average Precision and Recall Excl. MISC)",
    #format="latex",
    align=c("l", rep("c", nrow(.)-1)),
    table.attr = "style='width:30%;'"
    ) %>% 
  add_header_above(c(" " = 1, " " = 1, "All" = 2, "Men" = 2, "Women" = 2, "Unisex"=1)) %>% 
  collapse_rows() %>% 
  kable_classic(full_width = F, html_font = "Cambria") 
```


### Table 2: 
```{r}
collapse_rows_df(both_metrics_tb2, Model) %>% 
  kbl(
    booktabs=T, 
    col.names = c("Model", "Metric", "Majority", "Minority", "Majority", "Minority", "Majority", "Minority", "Majority"),
    caption = "NER scores for 'PER' Entity (Average Precision and Recall)",
    #format="latex",
    align=c("l", rep("c", nrow(.)-1)),
    table.attr = "style='width:30%;'"
    ) %>% 
  add_header_above(c(" " = 1, " " = 1, "All" = 2, "Men" = 2, "Women" = 2, "Unisex"=1)) %>% 
  collapse_rows() %>% 
  kable_classic(full_width = F, html_font = "Cambria") 
```


